Effect of quality-adjusted production function inputs

### 2015-06-19 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Now using CST models as the basis of comparison

# Tell whether to re-run the analysis to create the models.
# This file will use previously-saved models to generate
# the paper if set to FALSE, greatly reducing time.
# If something changes in the data or in the code that
# performs the analysis, set 
# rerunModels = TRUE
# once to re-generate the information needed to 
# create the paper.
# ********************************************************
rerunModels=FALSE
# ********************************************************
opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE,
  # Sets all table captions to the top.
  xtable.caption.placement="top"
)
@

@@

@@

We can test our hypothesis that quality-adjusting the factors of production
and including energy will both reduce the Solow residuals~($\lambda$)
and improve the fit to historical data
(thereby, reducing the fitting residuals, $r_i$)
by calculating 
$\Delta \lambda$ and $\Delta mse$, where
%
\begin{equation} \label{eq:delta_lambda}
  \Delta \lambda \equiv \left| \lambda \right| - \left| \lambda_{Unadjusted, \, Without \, energy} \right|
\end{equation}
%
and
%
\begin{equation} \label{eq:delta_mse}
  \Delta mse \equiv mse - mse_{Unadjusted, \, Without \, energy}.
\end{equation}
%
where~$N$ is the number of years of data.
On a percentage basis,
%
\begin{equation} \label{eq:delta_lambda_perc}
  \Delta \lambda \, [\%] = \frac{100 \, \Delta \lambda}{\left| \lambda_{Unadjusted, \, Without \, energy} \right|} 
            = \frac{100 \, \left| \lambda \right|}{\left| \lambda_{Unadjusted, \, Without \, energy} \right|} - 1
\end{equation}
%
When $\Delta \lambda$, $\Delta \lambda \, [\%]$, or $\Delta mse$ are negative, 
we observe reduction in the Solow residual ($\lambda$) or
the fitting residuals ($mse$)
relative to the unadjusted, no-energy condition.

@@

<<MyXTableFunction>>=
myXTable <- function(data, caption, label, digits=2){
  table <- xtable(data, caption=caption, label=label, digits=digits)
  print(table, 
        caption.placement="top", 
        size="\\tiny", 
        include.rownames=FALSE, 
        math.style.negative=TRUE)
}

@@

myXTable(filter(AllCoeffs, model=="CD") %>% select(Country, model, flavor, energy, Dlambda, Dmse),

@@

           select(Country, model, flavor, energy, nest, Dlambda, Dmse),

@@

<<DSRmse_CD_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Change in Solow Residuals~($\\Delta\\lambda$) and mean squared error~($\\Delta mse$) for the Cobb-Douglas model relative to the Unadjusted, Without energy condition.">>=
ggplot(data=filter(AllCoeffs, model=="CD")) +
  geom_point(mapping=aes(x=Dmse, y=Dlambda, color = energy)) +

@@

@@

<<percDSRmse_CD_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Percentage change in Solow Residuals~($\\Delta\\lambda \\, [\\%]$) and mean squared error~($\\Delta mse$) for the Cobb-Douglas model relative to the Unadjusted, Without energy condition.">>=
ggplot(data=filter(AllCoeffs, model=="CD")) +
  geom_point(mapping=aes(x=Dmse, y=percDlambda, color = energy)) +

@@

  scale_y_continuous(breaks=c(-50, 0, 50, 100)) +

@@

<<DSRmse_CES_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Change in Solow Residuals~($\\Delta\\lambda$) and mean squared error~($\\Delta mse$) for the CES modles relative to the Unadjusted, Without energy condition.">>=

@@

  geom_point(mapping=aes(x=Dmse, y=Dlambda, color = energy, shape=nest)) +

@@

<<percDSRmse_CES_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Percentage change in Solow Residuals~($\\Delta\\lambda \\, [\\%]$) and mean squared error~($\\Delta mse$) for the CES modles relative to the Unadjusted, Without energy condition.">>=

@@

  geom_point(mapping=aes(x=Dmse, y=percDlambda, color = energy, shape=nest)) +

@@

  scale_y_continuous(breaks=c(-50, -25, 0, 25)) +  

>>>>

# Tell whether to re-run the analysis to create the models.
# This file will use previously-saved models to generate
# the paper if set to FALSE, greatly reducing time.
# If something changes in the data or in the code that
# performs the analysis, set 
# rerunModels = TRUE
# once to re-generate the information needed to 
# create the paper.
# ********************************************************
rerunModels=FALSE
# ********************************************************
opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE,
  # Sets all table captions to the top.
  xtable.caption.placement="top"
)

# This function is used in many places below.
myXTable <- function(data, caption, label, digits=2){
  table <- xtable(data, caption=caption, label=label, digits=digits)
  print(table, 
        caption.placement="top", 
        size="\\tiny", 
        include.rownames=FALSE, 
        math.style.negative=TRUE)
}
@

@@

%++++++++++++++++++++++++++++++
\subsection{Does rejecting the cost-share theorem 
            decrease the mean squared error and/or the Solow residual?} 
\label{sec:cst_and_solow_residual}
%++++++++++++++++++++++++++++++

Our hypothesis is that rejecting the cost-share theorem~(CST) will decrease
both mean squared error~($mse$) and
the Solow residual~($\lambda$).
The mean squared error~($mse$) is expected to decrease 
because rejecting the CST adds one or more fitting parameters
to the model.
The Solow residual~($\lambda$) is expected to decrease 
because the factor shares for capital~($\alpha_1$) and labor~($\alpha_2 = 1 - \alpha_1$)
are free to float, 
allowing each factor of production 
to contribute optimally toward production and
bringing the model's prediction for economic output closer to 
historical output. 
In so doing, we expect less ``unexplained'' economic growth 
and a decreased Solow residual~($\lambda$).

Table~\ref{tab:Effect_of_CST_CD} shows the effect of the CST 
on fitted parameters for the Cobb-Douglas model.
%
<<Effect_of_CST_CD, results="asis">>=
myXTable(filter(AllCoeffs, model=="CD" & flavor=="Unadjusted" & energy=="Without energy") %>% 
           select(Country, cst, scale, lambda, alpha_1, alpha_2, mse),
         caption="Model parameters for Cobb-Douglas models with unadjusted factors of production, without energy.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
Table~\ref{tab:Effect_of_CST_CES} shows the effect of the CST 
on fitted parameters for the CES model.
%
<<Effect_of_CST_CES, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & flavor=="Unadjusted" & energy=="Without energy") %>% 
           select(Country, cst, gamma, lambda, alpha_1, alpha_2, sigma_1, mse),
         caption="Model parameters for CES models with unadjusted factors of production, without energy.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
For both Cobb-Douglas and CES, rejecting the Cost Share Theorem yields
decreased mean squared error~($mse$) and
smaller Solow residual ($\lambda$, on an absolute-value basis),
as expected.

***************

Note that in Table~\ref{tab:Effect_of_CST_CES}, we see $\sigma_1 = 0.5$ 
for UK when the cost share theorem is rejected. 
This even-numbered result is coming from the grid search in $\rho_1$. 
For an unknown reason, the gradient search from the best grid search point
fails.
I will investigate.

This result means that we have found a set of fitting coefficients
for this situation 
that is \emph{close} to providing the minimum possible $mse$,
but it may not be the \emph{exact} minimum.
Regardless, the point above (that $mse$ and $\lambda$ always decrease 
when rejecting the cost share theorem) remains valid.
Investigating and fixing this problem
will only serve to \emph{further} reduce $mse$ and $\lambda$ 
from the values reported in Table~\ref{tab:Effect_of_CST_CES}.

************

@@

We can test our hypothesis that 
quality-adjusting the factors of production
and including energy will both reduce the Solow residuals~($\lambda$)
and improve the fit to historical data
(thereby reducing the fitting residuals, $r_i$, and mean squared error, $mse$)
by calculating 
$\Delta \lambda$ and $\Delta mse$, where
%
\begin{equation} \label{eq:delta_lambda}
  \Delta \lambda \equiv \left| \lambda \right| - \left| \lambda_{Unadjusted, \, Without \, energy, \, CST} \right|
\end{equation}
%
and
%
\begin{equation} \label{eq:delta_mse}
  \Delta mse \equiv mse - mse_{Unadjusted, \, Without \, energy, \, CST}.
\end{equation}
%
where~$N$ is the number of years of data.
On a percentage basis,
%
\begin{equation} \label{eq:delta_lambda_perc}
  \Delta \lambda \, [\%] = \frac{100 \, \Delta \lambda}{\left| \lambda_{Unadjusted, \, Without \, energy, \, CST} \right|} 
            = \frac{100 \, \left| \lambda \right|}{\left| \lambda_{Unadjusted, \, Without \, energy, \, CST} \right|} - 1
\end{equation}
%
When $\Delta \lambda$, $\Delta \lambda \, [\%]$, or $\Delta mse$ are negative, 
we observe reduction in the Solow residual ($\lambda$) or
the fitting residuals ($mse$)
relative to the unadjusted, no-energy condition.

@@

@@

myXTable(filter(AllCoeffs, model=="CD") %>% select(Country, model, flavor, energy, cst, Dlambda, Dmse),

@@

           select(Country, model, flavor, energy, nest, cst, Dlambda, Dmse),

@@

<<DSRmse_CD_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Change in Solow Residuals~($\\Delta\\lambda$) and mean squared error~($\\Delta mse$) for the Cobb-Douglas model relative to the Unadjusted, Without energy, CST case.">>=
ggplot() +
  geom_point(data=filter(AllCoeffs, cst==FALSE & model=="CD"), mapping=aes(x=Dmse, y=Dlambda, color=energy)) +
  scale_shape(solid=FALSE) +
  guides(shape=FALSE) +
  geom_point(data=filter(AllCoeffs, cst==TRUE & model=="CD"), mapping=aes(x=Dmse, y=Dlambda, color=energy, shape=cst)) +

@@

  scale_x_continuous(breaks=seq(-0.005, 0, 0.005)) +  

@@

<<percDSRmse_CD_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Percentage change in Solow Residuals~($\\Delta\\lambda \\, [\\%]$) and mean squared error~($\\Delta mse$) for the Cobb-Douglas model relative to the Unadjusted, Without energy, CST case.">>=
ggplot() +
  geom_point(data=filter(AllCoeffs, cst==FALSE & model=="CD"), mapping=aes(x=Dmse, y=percDlambda, color=energy)) +
  scale_shape(solid=FALSE) +
  guides(shape=FALSE) +
  geom_point(data=filter(AllCoeffs, cst==TRUE & model=="CD"), mapping=aes(x=Dmse, y=percDlambda, color=energy, shape=cst)) +

@@

  scale_y_continuous(breaks=seq(-75, 25, 25)) +  

@@

<<DSRmse_CES_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Change in Solow Residuals~($\\Delta\\lambda$) and mean squared error~($\\Delta mse$) for the CES modles relative to the Unadjusted, Without energy, CST case.">>=

@@

  geom_point(mapping=aes(x=Dmse, y=Dlambda, color=energy, shape=nest)) +

@@

<<percDSRmse_CES_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Percentage change in Solow Residuals~($\\Delta\\lambda \\, [\\%]$) and mean squared error~($\\Delta mse$) for the CES modles relative to the Unadjusted, Without energy, CST case.">>=

@@

  geom_point(mapping=aes(x=Dmse, y=percDlambda, color=energy, shape=nest)) +

@@

  scale_y_continuous(breaks=seq(-75, 25, 25)) +  
